---
title: "US data - making clusters based on flyway and fixing dates"
author: "Tanin Rajamand"
date: "2024-08-10"
output:
  pdf_document: default
  html_document: default
---

This dataset is for the presentation on Sep 3, 2024. This presentation is dedicated to present the data to potential collaborators. 


before we begin:
```{r}
library(dplyr)
library(tibble)
library(ggplot2)
library(stringr)
library(RColorBrewer)
library(wesanderson)
library(lubridate)
library(readxl)
library(tidyverse)
```


lets load the dataset:
```{r}
print(getwd())
# Set the working directory (this is already set in your case)
setwd("C:/Users/tanin/OneDrive/Documents/Lab R/Cattle movement/Data Presentation")

# Load the dataset using just the file name since the working directory is already set correctly
Final_For_Fasta <- read_excel("Final For Fasta.xlsx")

```
lets look at the cols and keep the ones I want:
```{r}
colnames(Final_For_Fasta)
Final_For_Fasta <- subset(Final_For_Fasta, select = c(Isolate_Name, Isolate_Id, Location, Continent, Dom_Wild, Country, Year, Collection_Date, Host_category, Sequence))
```
let's subset the US:
```{r}
sep <- Final_For_Fasta %>%
  filter(Country == "United States")
```



now lets get the states:
```{r}
broken_loc <- strsplit(sep$Location, "/")
state <- sapply(broken_loc , function(x) x[3])

sep<- mutate(sep, State= state)
```


lets see the unique states:
```{r}
unique_states <- unique(sep$State)
unique_states
```
I saw an NA, lets see this first.
then I wil clean the white spaces and all

```{r}
na_states <- sep %>%
  filter(is.na(State))

na_states
```

after studying this, i noticed i can use isolate name to overwrite the state instead of NA.

I also noticed I have Unknown and USA in there as well

```{r}
sep <- sep %>%
  mutate(State = if_else(is.na(State),
                         sapply(strsplit(Isolate_Name, "/"), function(parts) {
                           if(length(parts) >= 3) parts[3] else NA
                         }),
                         State))
```


lets look at the unique ones and the NA:
```{r}

unique_states2 <- unique(sep$State)
unique_states2

na_states2 <- sep %>%
  filter(is.na(State))

na_states2
```

Let's clean it a bit:
start with white space
```{r}
sep$State <- trimws(sep$State)


unique_states3 <- unique(sep$State)
unique_states3
```
Found some problems here:
Deleware (should be Delaware)
Conneticut (should be Connecticut)
MD (abbreviation for Maryland)
Massachussetts (should be Massachusetts)
USA
Unknown

```{r}
specific_states <- c("Deleware", "Conneticut", "MD", "Massachussetts", "USA", "Unknown")
specific_rows <- sep %>% filter(State %in% specific_states)

print(specific_rows)
```

let's rename them:
```{r}
sep <- sep %>%
  mutate(State = case_when(
    State == "Deleware" ~ "Delaware",
    State == "Conneticut" ~ "Connecticut",
    State == "MD" ~ "Maryland",
    State == "Massachussetts" ~ "Massachusetts",
    State == "Unknown" ~ "USA",
    TRUE ~ State
  ))

unique_states4 <- unique(sep$State)
unique_states4

```


```{r}
sep <- sep %>%
  mutate(Flyway = case_when(
    State %in% c("Connecticut", "Delaware", "Florida", "Georgia", "Maine", "Maryland", 
                 "Massachusetts", "New Hampshire", "New Jersey", "New York", 
                 "North Carolina", "Pennsylvania", "Rhode Island", "South Carolina", 
                 "Vermont", "Virginia", "West Virginia") ~ "Atlantic Flyway",
    State %in% c("Alabama", "Arkansas", "Indiana", "Illinois", "Iowa", "Kentucky", 
                 "Louisiana", "Michigan", "Minnesota", "Mississippi", 
                 "Missouri", "Ohio", "Tennessee", "Wisconsin") ~ "Mississippi Flyway",
    State %in% c("Montana", "Wyoming", "Colorado", "New Mexico", "Texas", 
                 "Oklahoma", "Kansas", "Nebraska", "South Dakota", "North Dakota") ~ "Central Flyway",
    State %in% c("Alaska", "Arizona", "California", "Idaho", "Nevada", 
                 "Oregon", "Utah", "Washington", "Colorado", "Montana", 
                 "New Mexico", "Wyoming") ~ "Pacific Flyway",
    TRUE ~ "Other"
  ))

head(sep)

```

lets look at the others - i should only have the ones called US:

```{r}
other_states <- sep %>%
  filter(Flyway == "Other") %>%
  select(State, Flyway) %>%
  distinct()

print(other_states)

```
rows:
```{r}
other_states <- sep %>%
  filter(Flyway == "Other")

print(other_states)

```

great now i need to make sure the dates are ok and then make a fasta file to run more analysis. 

lets convert the dates:
```{r}
# Convert the Collection_Date column
sep <- sep %>%
  mutate(Collection_Date = ifelse(grepl("^[0-9]+$", Collection_Date),
                                  as.character(as.Date(as.numeric(Collection_Date), origin = "1899-12-30")),
                                  Collection_Date))

# Check for any NA values that resulted from the conversion
na_dates <- sep %>% filter(is.na(Collection_Date))

# Print out the problematic entries, if any
print(na_dates)
```


lets download the dataset:
```{r}
library(writexl)
write_xlsx(sep, "C:/Users/tanin/OneDrive/Documents/Lab R/Cattle movement/Data Presentation/US_Flyway_Clusters.xlsx")

```



