---
title: "Data Presentation"
author: "Tanin Rajamand"
date: "2024-08-08"
output:
  pdf_document: default
  html_document: default
---


This dataset is for the presentation on Sep 3, 2024. This presentation is dedicated to present the data to potential collaborators. 


before we begin:
```{r}
library(dplyr)
library(tibble)
library(ggplot2)
library(stringr)
library(RColorBrewer)
library(wesanderson)
library(lubridate)
library(readxl)
library(tidyverse)
```


lets load the dataset:
```{r}
print(getwd())
# Set the working directory (this is already set in your case)
setwd("C:/Users/tanin/OneDrive/Documents/Lab R/Cattle movement/Data Presentation")

# Load the dataset using just the file name since the working directory is already set correctly
Final_For_Fasta <- read_excel("Final For Fasta.xlsx")

```
lets look at the cols and keep the ones I want:
```{r}
colnames(Final_For_Fasta)
Final_For_Fasta <- subset(Final_For_Fasta, select = c(Isolate_Name, Isolate_Id, Location, Continent, Dom_Wild, Country, Year, Collection_Date, Host_category, Sequence))
```
let's subset the US:
```{r}
sep <- Final_For_Fasta %>%
  filter(Country == "United States")
```



now lets get the states:
```{r}
broken_loc <- strsplit(sep$Location, "/")
state <- sapply(broken_loc , function(x) x[3])

sep<- mutate(sep, State= state)
```


lets see the unique states:
```{r}
unique_states <- unique(sep$State)
unique_states
```
I saw an NA, lets see this first.
then I wil clean the white spaces and all

```{r}
na_states <- sep %>%
  filter(is.na(State))

na_states
```

after studying this, i noticed i can use isolate name to overwrite the state instead of NA.

I also noticed I have Unknown and USA in there as well

```{r}
sep <- sep %>%
  mutate(State = if_else(is.na(State),
                         sapply(strsplit(Isolate_Name, "/"), function(parts) {
                           if(length(parts) >= 3) parts[3] else NA
                         }),
                         State))
```


lets look at the unique ones and the NA:
```{r}

unique_states2 <- unique(sep$State)
unique_states2

na_states2 <- sep %>%
  filter(is.na(State))

na_states2
```

Let's clean it a bit:
start with white space
```{r}
sep$State <- trimws(sep$State)


unique_states3 <- unique(sep$State)
unique_states3
```
Found some problems here:
Deleware (should be Delaware)
Conneticut (should be Connecticut)
MD (abbreviation for Maryland)
Massachussetts (should be Massachusetts)
USA
Unknown

```{r}
specific_states <- c("Deleware", "Conneticut", "MD", "Massachussetts", "USA", "Unknown")
specific_rows <- sep %>% filter(State %in% specific_states)

print(specific_rows)
```

let's rename them:
```{r}
sep <- sep %>%
  mutate(State = case_when(
    State == "Deleware" ~ "Delaware",
    State == "Conneticut" ~ "Connecticut",
    State == "MD" ~ "Maryland",
    State == "Massachussetts" ~ "Massachusetts",
    State == "Unknown" ~ "USA",
    TRUE ~ State
  ))

unique_states4 <- unique(sep$State)
unique_states4

```

lets look at the logistics now 
lets start with the states
```{r}
ggplot(sep, aes(x=State, fill=State)) +
  geom_bar() +
  xlab("State") +
  ylab("Count") +
  ggtitle("Count of Different States") +
  geom_text(stat='count', aes(label=..count..), vjust=-0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

this wasnt enough, ill do this again in tableau at the end

now for dom wild
```{r}
ggplot(sep, aes(x=Dom_Wild, fill=Dom_Wild)) +
  geom_bar() +
  xlab("Dom_Wild") +
  ylab("Count") +
  ggtitle("Count of Different Dom_Wild Categories") +
  geom_text(stat='count', aes(label=..count..), vjust=-0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Host categories:
```{r}
ggplot(sep, aes(x = Host_category, fill = Host_category)) +
  geom_bar() +
  xlab("Host Category") +
  ylab("Count") +
  ggtitle("Count of Different Host Categories") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

look at pets in different
```{r}
pet_hosts <- sep %>%
  filter(Dom_Wild == "Pet")

# Create a bar plot for host categories within 'Pet'
ggplot(pet_hosts, aes(x = Host_category, fill = Host_category)) +
  geom_bar() +
  xlab("Host Category") +
  ylab("Count") +
  ggtitle("Count of Host Categories in Pets") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

where are pets?
```{r}
unique_states <- pet_hosts %>%
  pull(State) %>%
  unique()

print(unique_states)


ggplot(pet_hosts, aes(x = State, fill = State)) +
  geom_bar() +
  xlab("State") +
  ylab("Count") +
  ggtitle("Count of Pet Hosts by State") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

for humans:
```{r}
human_samples <- sep %>%
  filter(Dom_Wild == "Human")

unique_states_human <- human_samples %>%
  pull(State) %>%
  unique()

print(unique_states_human)

# Create a bar plot for states of 'Human' samples
ggplot(human_samples, aes(x = State, fill = State)) +
  geom_bar() +
  xlab("State") +
  ylab("Count") +
  ggtitle("Count of Human Samples by State") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

first look at wild bird categories and then where they are - this is going to be the pattern for all

```{r}
wild_bird_hosts <- sep %>%
  filter(Dom_Wild == "Wild Bird")

unique_host_categories <- wild_bird_hosts %>%
  pull(Host_category) %>%
  unique()

print(unique_host_categories)

ggplot(wild_bird_hosts, aes(x = Host_category, fill = Host_category)) +
  geom_bar() +
  xlab("Host Category") +
  ylab("Count") +
  ggtitle("Count of Host Categories in Wild Birds") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


now for wild bird:
```{r}
wild_bird_samples <- sep %>%
  filter(Dom_Wild == "Wild Bird")

unique_states_wild_bird <- wild_bird_samples %>%
  pull(State) %>%
  unique()

print(unique_states_wild_bird)

# Create a bar plot for states of 'Wild Bird' samples
ggplot(wild_bird_samples, aes(x = State, fill = State)) +
  geom_bar() +
  xlab("State") +
  ylab("Count") +
  ggtitle("Count of Wild Bird Samples by State") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

host categories for other bird
```{r}
other_bird_hosts <- sep %>%
  filter(Dom_Wild == "Other Bird")

unique_host_categories_other_bird <- other_bird_hosts %>%
  pull(Host_category) %>%
  unique()

print(unique_host_categories_other_bird)

ggplot(other_bird_hosts, aes(x = Host_category, fill = Host_category)) +
  geom_bar() +
  xlab("Host Category") +
  ylab("Count") +
  ggtitle("Count of Host Categories in Other Birds") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


location for Other Bird:
```{r}
other_bird_samples <- sep %>%
  filter(Dom_Wild == "Other Bird")

unique_states_other_bird <- other_bird_samples %>%
  pull(State) %>%
  unique()

print(unique_states_other_bird)

ggplot(other_bird_samples, aes(x = State, fill = State)) +
  geom_bar() +
  xlab("State") +
  ylab("Count") +
  ggtitle("Count of Other Bird Samples by State") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
host categories for livestock avian
```{r}
livestock_avian_hosts <- sep %>%
  filter(Dom_Wild == "Livestock Avian")

unique_host_categories_livestock_avian <- livestock_avian_hosts %>%
  pull(Host_category) %>%
  unique()

print(unique_host_categories_livestock_avian)

ggplot(livestock_avian_hosts, aes(x = Host_category, fill = Host_category)) +
  geom_bar() +
  xlab("Host Category") +
  ylab("Count") +
  ggtitle("Count of Host Categories in Livestock Avian") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```





location for Livestock Avian
```{r}
livestock_avian_samples <- sep %>%
  filter(Dom_Wild == "Livestock Avian")

unique_states_livestock_avian <- livestock_avian_samples %>%
  pull(State) %>%
  unique()

print(unique_states_livestock_avian)

ggplot(livestock_avian_samples, aes(x = State, fill = State)) +
  geom_bar() +
  xlab("State") +
  ylab("Count") +
  ggtitle("Count of Livestock Avian Samples by State") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

host categories for livestock mammal:
```{r}
livestock_mammal_hosts <- sep %>%
  filter(Dom_Wild == "Livestock Mammal")

unique_host_categories_livestock_mammal <- livestock_mammal_hosts %>%
  pull(Host_category) %>%
  unique()

print(unique_host_categories_livestock_mammal)

ggplot(livestock_mammal_hosts, aes(x = Host_category, fill = Host_category)) +
  geom_bar() +
  xlab("Host Category") +
  ylab("Count") +
  ggtitle("Count of Host Categories in Livestock Mammal") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


location for livestock mammal:
```{r}
livestock_mammal_samples <- sep %>%
  filter(Dom_Wild == "Livestock Mammal")

unique_states_livestock_mammal <- livestock_mammal_samples %>%
  pull(State) %>%
  unique()

print(unique_states_livestock_mammal)

ggplot(livestock_mammal_samples, aes(x = State, fill = State)) +
  geom_bar() +
  xlab("State") +
  ylab("Count") +
  ggtitle("Count of Livestock Mammal Samples by State") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

host categories for wild mammal:
```{r}
wild_mammal_hosts <- sep %>%
  filter(Dom_Wild == "Wild Mammal")

unique_host_categories_wild_mammal <- wild_mammal_hosts %>%
  pull(Host_category) %>%
  unique()

print(unique_host_categories_wild_mammal)

ggplot(wild_mammal_hosts, aes(x = Host_category, fill = Host_category)) +
  geom_bar() +
  xlab("Host Category") +
  ylab("Count") +
  ggtitle("Count of Host Categories in Wild Mammal") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


location for wild mammal:
```{r}
wild_mammal_samples <- sep %>%
  filter(Dom_Wild == "Wild Mammal")

unique_states_wild_mammal <- wild_mammal_samples %>%
  pull(State) %>%
  unique()

print(unique_states_wild_mammal)

ggplot(wild_mammal_samples, aes(x = State, fill = State)) +
  geom_bar() +
  xlab("State") +
  ylab("Count") +
  ggtitle("Count of Wild Mammal Samples by State") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

now lets get the counts:
```{r}
state_dom_wild_counts <- sep %>%
  group_by(State, Dom_Wild) %>%
  summarise(Count = n(), .groups = "drop")
```

download the counts:
```{r}
write.csv(state_dom_wild_counts, "state_dom_wild_counts.csv", row.names = FALSE)

```


now lets map them on tablue 
first we need to download this dataset though

```{r}
write.csv(sep, "sep.csv", row.names = FALSE)
```



i have some hope: 
lets see if i can make the map here








